### 2020-12-07 paper.tex
first commit

\documentclass[sigconf, review=true]{acmart}
\author{Jeroen-Niclas Trzaska}
\title{DB pro sem - isolation levels}
\begin{document}

\begin{abstract}
    abstract
\end{abstract}
\maketitle
\section{Intro}
\section{Phemomena Definitions}
We're now going to introduce the strengthened ANSI phenomena definitions taken from (cite here) which
are based on the broad interpretation of the ANSI standard (cite).

\subsection{P0 - Dirty write}
Lets start by defining the first phenomena, describing two consecutive writes by two differnt
transactions leading to an unclear value for x when the first transaction aborts.

\begin{example}
    Alice deposits 10€ into a bank account X.
    Bob than deposits a further 20€ into the same account.
    To do so he reads the value of X and then adds 20 and writes it back to the account.
    Allice than redacts her transaction.
    Thus the amount of monney in account X is unclear.
\end{example}

\subsection{P1 - Dirty read}
For the second phenomenon we're going to look at reading incosistent data.
This can be a problem when retrieving information while another transaction
that writes data is being executed.
\begin{example}
    Alice transfers 40€ form account x to account y.
    To do so she reads the value of x subtracts the 40€.
    Afterwards, Bob reads the value of x (10€) and y(15€) resulting in a sum of 25€
    Alice then reads the value of y (15€) and then adds the 40€ to it resulting in a 55€ sum.
\end{example}

\subsection{P2 - Fuzzy read}
The next phenomenon is similar to P1 as it to has to do with read inconsitencies.
Though this time performing the two reads of the reading transaction without interuption would lead
to the correct data, so the read is non repeatable.
\begin{example}
    Bob starts to read the sum by reading x and gets interrupted by Alice who then transfers 40€ from x to y after this is completed Bob reads the value of y.
    Thus the sum is 100 for Alice and 140 for Bob
    In contrast to P1 a second read of x by Bob would have returned the correct sum.
\end{example}
\subsection{P3 - Phantom}

This phonomenon is only applicable to transactions using predicates and aplies to deletions, insets, and upadates.
\begin{example}
    Alice reads the predicate blue cars. Bob then adds a new blue car to the database and increases
    the count of blue cars. Alice then reads the count of blue cars. This leads to a phantom item
    ( Alice can’t see Bob's car but knows of its existence due to the count).
\end{example}

\section{ANSI isolation levels}
We define the isolation levels by the phenomena prohibited by it and we define
Degree 0 isolation to be the isolation level that doesn’t prohibit any abnormal
behaviour. Furthermore, we define short duration locks as locks that are held
while a transaction reads or writes an item, in contrast to long-duration locks which are
held until a transaction commits or aborts.

\subsection{Read uncommitted}
We define read uncommitted to be the isolation level that prohibits the phenomenon
P0, dirty write, this can be accomplished by placing long-duration write locks when modifying a data item.

\subsection{Read committed}
Read committed is defined as the isolation level where P1, dirty read, is prohibited.
To accomplish this the use of short duration read locks, as well as long write locks,  is
necessary.

\subsection{Repeatable read}
For Repeatable read, the phenomenon P2, fuzzy read, has to be prohibited.
This requires long duration write locks as well as long duration read locks on
items and short ones for Predicates.\\
The difference to read committed is that read committed only guarantees that
the data read was committed at the time of reading (no dirty reads), whereas
repeatable read also guarantees that the data will not change before the
transaction finishes either with a commit or abort.

\subsection{Serializable}
For serializability P3 needs to be prohibited, which can be done by placing long locks
on reads and writes.

\section{Motivation}
The motivation behind these theoretical analyses is to better understand the
behaviour of different systems when presented with multi-item dependencies.

\section{Cursor stability}
For the cursor stability, it is useful to define a fourth phenomenon P4 (lost update).
\subsection{ P4  - lost update}

Alice reads x to be 100€. Bob then reads the same, adds 20€ and writes 120 to x and commits.
Alice then subtracts 40€ from x and writes 60€ to x and commits. Resulting in the loss of Bob's update.
\subsection{Implications}
The phenomena define above can be prohibited when a read lock is placed on x (position of the cursor) when
read and placing a long write lock on the items row when modified.
The read lock is removed when the cursor moves (e.g. reading a new cell),
while the write lock only gets removed after the transaction commits or aborts.
As Cursor stability is stronger then read commited while being weaker then repeatable read it
is used instead of the weaker repeatable read by some database systems. This behaviour is allowed by the ansi standard
as it reqiures a minimum set of guarantees.
\subsection{Multi cursor}
The system described above could also be extended with the use of multiple cursors thus
parleying the effect of repeatable read isolation as long as the transactions dont acces more items
then the system has cursors. But obiously this isn't a general, or practical solution to the phenomenon P2.

\section{Snapshot Isolation}
All of the above-mentioned systems suggest lock-based implementations. We will now look at
snapshot isolation which in contrast uses multi-versioning.
In snapshot isolation, each transaction gets its own snapshot of committed data at the start.
Thus reads are never blocked, while writes are performed on the given snapshot as to not disturb
other transactions and allow reads of the updated data only by the same transaction (unless committed).
For the committing procedure, a second (commit) timestamp is assigned, said timestamp is larger than every
other assigned time stamp in the system. If another transaction A already modified data the commit of
transaction B fails, leading to an abort if said data was modified by transaction B.
\begin{example}
    (first commit wins): \\
    Alice reads x and y to be 50€. Bob then reads the same data, while getting his own snapshot
    to operate on (newer timestamp than Alice). Alice then sets y to be -40€ and commits,
    resulting in a sum of 10€. Bob then sets y to be 30€ and tries to commit. As Alice already
    modified x Bob’s transaction will abort.
\end{example}
\subsection{Strength}
Snapshot isolation is comparable in strength to repeatable read

\end{document}

### 2020-12-07 paper.tex
updated paper

@@
 \documentclass[sigconf, review=true]{acmart}
+\usepackage[english]{babel}
 \author{Jeroen-Niclas Trzaska}
 \title{DB pro sem - isolation levels}
+\affiliation{%
+   \institution{TU Dresden}
+   \city{Dresden}
+   \state{Saxony}
+   \country{Germany}}
+\email{jeroen@trzaska.xyz}
+
+\citestyle{acmauthoryear}
 \begin{document}
@@
 \begin{abstract}
-    abstract
+
+    This paper is going to look at the strengthend ANSI definition poropositions by \cite{Adya_Liskov_O_Neil_2000} as well as the coresponding isolation levels,
+    comparing them in terms of strength. We will also take a quick look at non lock based implementations
+    and the problems the ANSI standart has in regards to such implementations as shown in \cite{Berenson_Bernstein_Gray_Melton_O_Neil_O_Neil_1995}.
 \end{abstract}
 \maketitle
-\section{Intro}
 \section{Phemomena Definitions}
-We're now going to introduce the strengthened ANSI phenomena definitions taken from (cite here) which
-are based on the broad interpretation of the ANSI standard (cite).
+We're now going to introduce the strengthened ANSI phenomena definitions as described in \cite{Adya_Liskov_O_Neil_2000}  which
+are based on the broad interpretation of the ANSI standard. These definitions are inherinty stronger then their ANSI counter part,
+as they rule out other Phenomena not define in the following paragraph.\\
+One might assume that ANSI had these strengthened definitions in mind when specifing them as lock based implementations
+are by far the most common systems.
 
 \subsection{P0 - Dirty write}
-Lets start by defining the first phenomena, describing two consecutive writes by two differnt
+An example for the above mentioned problem with ANSI strength is that while a locking implementation of read uncommitted
+prevents the following phenomenon, a non locking system would not do so.
+So lets start by defining the first phenomena, describing two consecutive writes by two differnt
 transactions leading to an unclear value for x when the first transaction aborts.
@@
 \subsection{P3 - Phantom}
-
+Once again when looking at the ANSI specification one can see that it only disallowes inserts on the serializable
+isolation level while the following definition of P3 prohibits deletions, inserts, and updates.
 This phonomenon is only applicable to transactions using predicates and aplies to deletions, insets, and upadates.
@@
 P0, dirty write, this can be accomplished by placing long-duration write locks when modifying a data item.
+Thus only allowing one transaction at a time to modify a given value.
 
@@
 To accomplish this the use of short duration read locks, as well as long write locks, is
-necessary.
+necessary. Thus disallowing a write on data that was just read by another transaction.
 
@@
 behaviour of different systems when presented with multi-item dependencies.
+Such dependencies include but aren't limited to constrains between two items.
+
+Let's say that Alice has two acoounts at a bank. She has 20€ on each of them now she whithdraws 30€ from acount A and another 20€ form account B.
+The looking at each of the transactions we can see that the constrain $A+B \geq 0$ is met.
+But when executed simultaniously the resulting sum can still be negative.
+\begin{example}
+    Alice reads 20€ to be in both accounts and Bob does the same. Alice then withdraws 30€ from account A and writes -10€ into it and commits, the constrain is still met.
+    Bob then withdraws 20€ from account and , as his transaction read both acounts to have 20€, he writes 0€  to B. This now violates the constrain $A+B \geq 0$.
+\end{example}
 
@@
 \subsection{ P4  - lost update}
-
+This phenomenon describes the loss of an update by a transaction due to it being imeadiatly overwritten
+by a second transaction.
+\begin{example}
     Alice reads x to be 100€. Bob then reads the same, adds 20€ and writes 120 to x and commits.
     Alice then subtracts 40€ from x and writes 60€ to x and commits. Resulting in the loss of Bob's update.
+\end{example}
 \subsection{Implications}
@@
 parleying the effect of repeatable read isolation as long as the transactions dont acces more items
-then the system has cursors. But obiously this isn't a general, or practical solution to the phenomenon P2.
+then the system has cursors. But obiously this isn't a general, or even practical solution to the phenomenon P2.
 
@@
 other assigned time stamp in the system. If another transaction A already modified data the commit of
+\subsection{Strength}
+Snapshot isolation is comparable in strength to repeatable read as both prohibit P0, P1, and P2.
 transaction B fails, leading to an abort if said data was modified by transaction B.
+\subsection{P5B - Write skew}
+To take a closer look at the difference to repeatable read it is usefull to define a new phenomenon.
 \begin{example}
@@
 \end{example}
-\subsection{Strength}
-Snapshot isolation is comparable in strength to repeatable read
+\subsection{Comparison with repeatable read}
+\subsection{Comparison to ANSI}
+When looking at the weak interpretation of the ANSI specification as in \cite{Adya_Liskov_O_Neil_2000}
+one can see that Snapshot isolation prohibits all three anomlies A1, A2, and A3 making it ANOMALY SERIALIZABLE.
+To show this we will now take a look at A3 as we have already shown above that Snapshot isolation prohibits P1 and P2.
+\subsection{A3 - Anomaly Phantom}
+This is the strict interpretation of the Phatom anomaly. The difference to P3 is the position of the
+transactions comit is not indiferent.
+\begin{example}
+    Alice reads the predicate blue cars. Bob then adds a new blue car to the database and increases
+    the count of blue cars. Bob commits.  Alice then reads the count of blue cars. This leads to a phantom item
+    ( Alice can’t see Bob's car but knows of its existence due to the count).
+\end{example}
+Thus A3 only opperates on commited data while P3 allows for transactions to still be in prosses.
+\section{Conclusion}
+\bibliographystyle{plainnat}
+\bibliography{refs}
 

### 2020-12-13 paper.tex
further updates to formulations

@@
 \usepackage[english]{babel}
+\usepackage{graphicx}
+\graphicspath{ {./images/} }
 \author{Jeroen-Niclas Trzaska}
 \title{DB pro sem - isolation levels}
+\acmDOI{}
+\acmISBN{}
 \affiliation{%
@@
     comparing them in terms of strength. We will also take a quick look at non lock based implementations
-    and the problems the ANSI standart has in regards to such implementations as shown in \cite{Berenson_Bernstein_Gray_Melton_O_Neil_O_Neil_1995}.
+    and the problems the ANSI standard has in regards to such implementations as shown in \cite{Berenson_Bernstein_Gray_Melton_O_Neil_O_Neil_1995}.
 \end{abstract}
 \maketitle
+
+\section{Motivation}
+The motivation behind these analyses of the ANSI standard is to better define and examine the
+behaviour of different database implementation aproaches when presented with multi-item dependencies.
+This is important as the ANSI standard was created when locking implementations where the norm and thus multi version systems weren't
+taken into account creating it leading to a behaviour not in line with ANSI.
+These issues include but aren't limited to dependency constrains between two items.
+
+Let's say that Alice has two acoounts at a bank. She has 20€ on each of them now she whithdraws 30€ from acount A and another 20€ form account B.
+The looking at each of the transactions we can see that the constrain $A+B \geq 0$ is met.
+But when executed simultaniously the resulting sum can still be negative. To do so
+she asked Bob to help her.
+\begin{example}
+    Alice reads 20€ to be in both accounts and Bob does the same. Alice then withdraws 30€ from account A and writes -10€ into it and commits, the constrain is still met.
+    Bob then withdraws 20€ from account and, as his transaction read both acounts to have 20€, he writes 0€  to B. This now violates the constrain $A+B \geq 0$.
+\end{example}
 \section{Phemomena Definitions}
@@
 are based on the broad interpretation of the ANSI standard. These definitions are inherinty stronger then their ANSI counter part,
-as they rule out other Phenomena not define in the following paragraph.\\
-One might assume that ANSI had these strengthened definitions in mind when specifing them as lock based implementations
-are by far the most common systems.
+as they rule out other Phenomena not defined in the following paragraph. This is due to the fact that, as mentioned above, the ANSI standard
+was created with locking implementations in mind. Leading to definitions that would allow phenomena to apear in the database history when using non lock based systems
+while lock based systems using the same interpretation wouldn't.
 
@@
 prevents the following phenomenon, a non locking system would not do so.
-So lets start by defining the first phenomena, describing two consecutive writes by two differnt
-transactions leading to an unclear value for x when the first transaction aborts.
+So lets start by defining the first phenomena, which describes two consecutive writes by two differnt
+transactions leading to an unclear value for X when the first transaction aborts.
+
+\includegraphics[width=8cm]{P0}
 
@@
     Bob than deposits a further 20€ into the same account.
-    To do so he reads the value of X and then adds 20 and writes it back to the account.
+    To do so he reads the value of X and then adds 20€ and writes it back to the account X.
     Allice than redacts her transaction.
@@
 This can be a problem when retrieving information while another transaction
-that writes data is being executed.
+that writes data is being executed. An issue that can arive due to this phenomena is
+that the given transaction reading the inconsistent data might continue to work with this
+incorrect data and writing wrong data back to the database, corrupting it in the process.
+
+\includegraphics[width=8cm]{P1}
 \begin{example}
     Alice transfers 40€ form account x to account y.
-    To do so she reads the value of x subtracts the 40€.
+    To do so she reads the value of x subtracts the 40€ writing 10€ to value x.
     Afterwards, Bob reads the value of x (10€) and y(15€) resulting in a sum of 25€
-    Alice then reads the value of y (15€) and then adds the 40€ to it resulting in a 55€ sum.
+    Alice then reads the value of y (15€) and then adds the 40€ to it resulting in a 55€ sum for y
+    and a total of 65€.
 \end{example}
@@
 Though this time performing the two reads of the reading transaction without interuption would lead
-to the correct data, so the read is non repeatable.
+to the correct data, so the read is non repeatable. This is an issue as the transaction doesn't know that
+it was interupted and assumes that the data in the database hasn't changed. Thus it opperates under a false pretense
+and might corrupt the database when writing data back to it.
+
+\includegraphics[width=8cm]{P2}
 \begin{example}
     Bob starts to read the sum by reading x and gets interrupted by Alice who then transfers 40€ from x to y after this is completed Bob reads the value of y.
-    Thus the sum is 100 for Alice and 140 for Bob
+    Thus the sum is 100€ for Alice and 140€ for Bob.
     In contrast to P1 a second read of x by Bob would have returned the correct sum.
@@
 Once again when looking at the ANSI specification one can see that it only disallowes inserts on the serializable
-isolation level while the following definition of P3 prohibits deletions, inserts, and updates.
+isolation level while the following definition of P3 prohibits deletions, inserts, and updates, making it stronger than A3.
 This phonomenon is only applicable to transactions using predicates and aplies to deletions, insets, and upadates.
+
+\includegraphics[width=8cm]{P3}
 \begin{example}
@@
     the count of blue cars. Alice then reads the count of blue cars. This leads to a phantom item
-    ( Alice can’t see Bob's car but knows of its existence due to the count).
+    (Alice can’t see Bob's car but knows of its existence due to the increased count).
 \end{example}
+Not prohibiting this phenomenon can cause inconsistent predicate reads and in the worst case phantom entries in the
+database as the reading transaction could, after reading a phantom object, either assume that it doesn't exist or treat it like an actual
+object, but both would lead to undefined behaviour of the transaction.
 
@@
 held until a transaction commits or aborts.
+These definition can then be used to clasify different database system implementations in
+regards to their serializability.
 
@@
 To accomplish this the use of short duration read locks, as well as long write locks, is
-necessary. Thus disallowing a write on data that was just read by another transaction.
+necessary. Thus disallowing a write on data that was just read by another transaction as
+well as disallowing overwriting of an item already written by another, not finished, transaction.
 
@@
 For serializability P3 needs to be prohibited, which can be done by placing long locks
-on reads and writes.
-
-\section{Motivation}
-The motivation behind these theoretical analyses is to better understand the
-behaviour of different systems when presented with multi-item dependencies.
-Such dependencies include but aren't limited to constrains between two items.
-
-Let's say that Alice has two acoounts at a bank. She has 20€ on each of them now she whithdraws 30€ from acount A and another 20€ form account B.
-The looking at each of the transactions we can see that the constrain $A+B \geq 0$ is met.
-But when executed simultaniously the resulting sum can still be negative.
-\begin{example}
-    Alice reads 20€ to be in both accounts and Bob does the same. Alice then withdraws 30€ from account A and writes -10€ into it and commits, the constrain is still met.
-    Bob then withdraws 20€ from account and , as his transaction read both acounts to have 20€, he writes 0€  to B. This now violates the constrain $A+B \geq 0$.
-\end{example}
+on reads and writes. This isolation level is stronger then it's ANSI counter part due to the
+prohibition of prdicate updates and deletions as defined in P3, making it the only isolation level
+outside of the original ANSI ioslation level specifications.
 
 \section{Cursor stability}
-For the cursor stability, it is useful to define a fourth phenomenon P4 (lost update).
+For the cursor stability, it is useful to define a fourth phenomenon P4 (lost update), as this isolation level
+was designed to solve the problem of lost updates.
 \subsection{ P4 - lost update}
-This phenomenon describes the loss of an update by a transaction due to it being imeadiatly overwritten
+This phenomenon describes the loss of an update by a transaction due to it imeadiatly being overwritten
 by a second transaction.
+\includegraphics[width=8cm]{P4}
 \begin{example}
-    Alice reads x to be 100€. Bob then reads the same, adds 20€ and writes 120 to x and commits.
+    Alice reads x to be 100€. Bob then reads the same, adds 20€ and writes 120€ to x and commits.
     Alice then subtracts 40€ from x and writes 60€ to x and commits. Resulting in the loss of Bob's update.
-\end{example}
     \subsection{Implications}
@@
     As Cursor stability is stronger then read commited while being weaker then repeatable read it
-is used instead of the weaker repeatable read by some database systems. This behaviour is allowed by the ansi standard
-as it reqiures a minimum set of guarantees.
+    is used instead of the weaker repeatable read by some database systems. This behaviour is allowed by the ANSI standard
+    as it defines a minimum set of guarantees.
+\end{example}
+\subsection{Stength}
+When looking at our isiolation levels defined in section 2, we can see that P4 is prohibited by repeatable read,
+as long duration locks on read and write will also disallow the overwriting of an item by another transaction.
+It is thus weaker than repeatable read as prohibiting P4 doesn't prohibit P3.
+Looking at read commited one can reason that one transaction A can read X, then be interupted by transaction B which updates
+item X and commits, and transaction A then continues to write to X.
+Thus makin Cursor stability stronger than the strngthened read commited.
 \subsection{Multi cursor}
-The system described above could also be extended with the use of multiple cursors thus
-parleying the effect of repeatable read isolation as long as the transactions dont acces more items
-then the system has cursors. But obiously this isn't a general, or even practical solution to the phenomenon P2.
-
+While the Cursor stability based system described above could also be extended with the use of multiple cursors, thus
+parleying the effect of repeatable read isolation as long as each transaction doesn't acces more items
+then the system has cursors. But obiously this isn't a general, or even practical solution to the phenomenon P2 and
+although it might have real life usecases it won't be discussed further in this paper.
 \section{Snapshot Isolation}
-All of the above-mentioned systems suggest lock-based implementations. We will now look at
-snapshot isolation which in contrast uses multi-versioning.
+All of the above-mentioned systems suggest lock-based implementations, in contrast we will now look at
+snapshot isolation which uses multi-versioning.
 In snapshot isolation, each transaction gets its own snapshot of committed data at the start.
@@
 For the committing procedure, a second (commit) timestamp is assigned, said timestamp is larger than every
-other assigned time stamp in the system. If another transaction A already modified data the commit of
-\subsection{Strength}
-Snapshot isolation is comparable in strength to repeatable read as both prohibit P0, P1, and P2.
-transaction B fails, leading to an abort if said data was modified by transaction B.
+other assigned time stamp in the system. If another transaction A already modified data of the commiting
+transaction B it's commit fails, leading to an abort.
 \subsection{P5B - Write skew}
 To take a closer look at the difference to repeatable read it is usefull to define a new phenomenon.
+
+\includegraphics[width=8cm]{P5}
 \begin{example}
@@
 \end{example}
+
 \subsection{Comparison with repeatable read}
+In comparison to snapshot isolation repeatable read prohibits P5B while snapshot isolation prohibits some cases of
+P3.
+\subsection{Strength}
+Snapshot isolation is comparable in strength to repeatable read as both prohibit P0, P1, and P2.
 \subsection{Comparison to ANSI}
@@
     the count of blue cars. Bob commits.  Alice then reads the count of blue cars. This leads to a phantom item
-    ( Alice can’t see Bob's car but knows of its existence due to the count).
+    (Alice can’t see Bob's car but knows of its existence due to the increased count).
 \end{example}
-Thus A3 only opperates on commited data while P3 allows for transactions to still be in prosses.
+Thus A3 only opperates on commited data while P3 accounts for transactions to still be in prosses.
 \section{Conclusion}
+\includegraphics[width=8cm, height=8cm]{iso_lvl_dia.png}
+
+In conclusion one can see that the ANSI specification is inherintly flawed whwn talking
+about multi-version histories and non locking database systems, as it failed to consider
+these systems when defining the phenomena.
 \bibliographystyle{plainnat}

### 2020-12-13 paper.tex
updated paper

@@
 P0, dirty write, this can be accomplished by placing long-duration write locks when modifying a data item.
-Thus only allowing one transaction at a time to modify a given value.
+This only allows one transaction at a time to write to a given value and requires all others to wait for a commit or abort.
 
@@
 \begin{example}
-    (first commit wins): \\
-    Alice reads x and y to be 50€. Bob then reads the same data, while getting his own snapshot
-    to operate on (newer timestamp than Alice). Alice then sets y to be -40€ and commits,
-    resulting in a sum of 10€. Bob then sets y to be 30€ and tries to commit. As Alice already
-    modified x Bob’s transaction will abort.
+    (first commit wins):
+    Constrain: $ X+Y \geq  50$
+    Alice reads x to be 10€ and y to be 50€. Bob then reads the same data, while getting his own snapshot
+    to operate on (newer timestamp than Alice). Alice then sets x to be 5€,
+    resulting in a sum of 55€. Bob then sets y to be 40€. Allice commits and Bob then tries to commit.
 \end{example}
+This would be prohibited by repeatable read but is allowed in snapshot isolation.
+\subsection{Implications}
+When looking at the phenomena defined above one can se that prohibiting it with a locked based system
+would be the same as prohibiting P3 as a lost update can onl be prohibited by placing long locks.
+But when using a multi version system instead proscribing P5B isn't equal to proscribing P3.
 
+The example in 5.1 is allowed by Snapshot isolation as it does not check for constrains and the transaction
+doesn't modify the same item (x and y respectively). Repeatable read on the other hand would disallow such
+behaviour as it places long read and write locks on the items.
+P5B is thus usefull to differntiate between Snapshot isolation and repeatable read.
 \subsection{Comparison with repeatable read}
-In comparison to snapshot isolation repeatable read prohibits P5B while snapshot isolation prohibits some cases of
-P3.
+As staded above repeatable read prohibits P5B while snapshot isolation prohibits some cases of
+P3 and both prohibit P0, P1, and P2. Although the argument above is made with a locking implementation
+it is general as the phenomena defined were strengthened to match the locking behaviour.
 \subsection{Strength}
 Snapshot isolation is comparable in strength to repeatable read as both prohibit P0, P1, and P2.
-\subsection{Comparison to ANSI}
-When looking at the weak interpretation of the ANSI specification as in \cite{Adya_Liskov_O_Neil_2000}
-one can see that Snapshot isolation prohibits all three anomlies A1, A2, and A3 making it ANOMALY SERIALIZABLE.
-To show this we will now take a look at A3 as we have already shown above that Snapshot isolation prohibits P1 and P2.
+To further investigate their difference we have to take a look at the original ANSI specification of serializable,
+as it prohibits A3, which will be defined in the following section.
 \subsection{A3 - Anomaly Phantom}
 This is the strict interpretation of the Phatom anomaly. The difference to P3 is the position of the
-transactions comit is not indiferent.
+transactions commit is not indiferent for A3.
 \begin{example}
@@
 \end{example}
-Thus A3 only opperates on commited data while P3 accounts for transactions to still be in prosses.
+Thus A3 only opperates on commited data while P3 accounts for transactions to still be in prosses,
+making A3 a weaker phenomenon than P3.
+\subsection{Comparison to ANSI}
+When looking at the weak interpretation of the ANSI specification as in \cite{Adya_Liskov_O_Neil_2000}
+one can see that Snapshot isolation prohibits all three anomlies A1, A2, and A3 making it ANOMALY SERIALIZABLE.
+To show this we will now take a look at A3 as we have already shown above that Snapshot isolation prohibits P1 and P2.
+As the definition of A3 states the importance of commit sequence it is easy to see that Snapshot isolation with a first commit winns
+stategy precluders A3, as a already altered item will lead to an abort.
+When looking at a first write wins the same is true as still only one modified value will
+be written to the database and the other one will be discarded by an abort. It thus only changes
+which transaction winns in comparison to first commit wins.
+\subsection{Comparison to Serializable}
+In comparison to the serializable isolation level defined in 3.4 we can say that Snapshot isiolation is
+weaker as serializable precluders P5B as well as P3, which as shown in example 5.1 above doesn"t hold true
+for Snapshot isiolation.
 \section{Conclusion}
@@
 
-In conclusion one can see that the ANSI specification is inherintly flawed whwn talking
+In conclusion one can see that the ANSI specification is inherintly flawed when talking
 about multi-version histories and non locking database systems, as it failed to consider
-these systems when defining the phenomena.
+these systems when defining the phenomena. As this will inevitably lead to different interpretations
+and implementations it might be advisable to update the ANSI standard
+according to the definitions given in \cite{Adya_Liskov_O_Neil_2000} and \cite{Berenson_Bernstein_Gray_Melton_O_Neil_O_Neil_1995}
+to ensure the correct behaviour of non locking implementations of ANSI SQL.
 \bibliographystyle{plainnat}

### 2020-12-13 paper.tex
fixed spelling

--- 2020-12-13 paper.tex
@@
 
-    This paper is going to look at the strengthend ANSI definition poropositions by \cite{Adya_Liskov_O_Neil_2000} as well as the coresponding isolation levels,
+    This paper is going to look at the strengthened ANSI definition propositions by \cite{Adya_Liskov_O_Neil_2000} as well as the corresponding isolation levels,
     comparing them in terms of strength. We will also take a quick look at non lock based implementations
@@
 The motivation behind these analyses of the ANSI standard is to better define and examine the
-behaviour of different database implementation aproaches when presented with multi-item dependencies.
+behavior of different database implementation approaches when presented with multi-item dependencies.
 This is important as the ANSI standard was created when locking implementations where the norm and thus multi version systems weren't
-taken into account creating it leading to a behaviour not in line with ANSI.
+taken into account creating it leading to a behavior not in line with ANSI.
 These issues include but aren't limited to dependency constrains between two items.
 
-Let's say that Alice has two acoounts at a bank. She has 20€ on each of them now she whithdraws 30€ from acount A and another 20€ form account B.
+Let's say that Alice has two accounts at a bank. She has 20€ on each of them now she withdraws 30€ from account A and another 20€ form account B.
 The looking at each of the transactions we can see that the constrain $A+B \geq 0$ is met.
-But when executed simultaniously the resulting sum can still be negative. To do so
+But when executed simultaneously the resulting sum can still be negative. To do so
 she asked Bob to help her.
@@
     Alice reads 20€ to be in both accounts and Bob does the same. Alice then withdraws 30€ from account A and writes -10€ into it and commits, the constrain is still met.
-    Bob then withdraws 20€ from account and, as his transaction read both acounts to have 20€, he writes 0€  to B. This now violates the constrain $A+B \geq 0$.
+    Bob then withdraws 20€ from account and, as his transaction read both accounts to have 20€, he writes 0€  to B. This now violates the constrain $A+B \geq 0$.
 \end{example}
-\section{Phemomena Definitions}
+\section{Phenomena Definitions}
 We're now going to introduce the strengthened ANSI phenomena definitions as described in \cite{Adya_Liskov_O_Neil_2000}  which
-are based on the broad interpretation of the ANSI standard. These definitions are inherinty stronger then their ANSI counter part,
+are based on the broad interpretation of the ANSI standard. These definitions are inherently stronger then their ANSI counter part,
 as they rule out other Phenomena not defined in the following paragraph. This is due to the fact that, as mentioned above, the ANSI standard
-was created with locking implementations in mind. Leading to definitions that would allow phenomena to apear in the database history when using non lock based systems
+was created with locking implementations in mind. Leading to definitions that would allow phenomena to appear in the database history when using non lock based systems
 while lock based systems using the same interpretation wouldn't.
@@
 prevents the following phenomenon, a non locking system would not do so.
-So lets start by defining the first phenomena, which describes two consecutive writes by two differnt
+So lets start by defining the first phenomena, which describes two consecutive writes by two different
 transactions leading to an unclear value for X when the first transaction aborts.
@@
     To do so he reads the value of X and then adds 20€ and writes it back to the account X.
-    Allice than redacts her transaction.
-    Thus the amount of monney in account X is unclear.
+    Alice than redacts her transaction.
+    Thus the amount of money in account X is unclear.
 \end{example}
@@
 \subsection{P1 - Dirty read}
-For the second phenomenon we're going to look at reading incosistent data.
+For the second phenomenon we're going to look at reading inconsistent data.
 This can be a problem when retrieving information while another transaction
-that writes data is being executed. An issue that can arive due to this phenomena is
+that writes data is being executed. An issue that can arrive due to this phenomena is
 that the given transaction reading the inconsistent data might continue to work with this
@@
     To do so she reads the value of x subtracts the 40€ writing 10€ to value x.
-    Afterwards, Bob reads the value of x (10€) and y(15€) resulting in a sum of 25€
+    Afterwards, Bob reads the value of x (10€) and y (15€) resulting in a sum of 25€
     Alice then reads the value of y (15€) and then adds the 40€ to it resulting in a 55€ sum for y
@@
 \subsection{P2 - Fuzzy read}
-The next phenomenon is similar to P1 as it to has to do with read inconsitencies.
-Though this time performing the two reads of the reading transaction without interuption would lead
+The next phenomenon is similar to P1 as it to has to do with read inconsistencies.
+Though this time performing the two reads of the reading transaction without interruption would lead
 to the correct data, so the read is non repeatable. This is an issue as the transaction doesn't know that
-it was interupted and assumes that the data in the database hasn't changed. Thus it opperates under a false pretense
+it was interrupted and assumes that the data in the database hasn't changed. Thus it operates under a false pretense
 and might corrupt the database when writing data back to it.
@@
 \subsection{P3 - Phantom}
-Once again when looking at the ANSI specification one can see that it only disallowes inserts on the serializable
+Once again when looking at the ANSI specification one can see that it only disallows inserts on the serializable
 isolation level while the following definition of P3 prohibits deletions, inserts, and updates, making it stronger than A3.
-This phonomenon is only applicable to transactions using predicates and aplies to deletions, insets, and upadates.
+This phenomenon is only applicable to transactions using predicates and applies to deletions, insets, and updates.
 
@@
 database as the reading transaction could, after reading a phantom object, either assume that it doesn't exist or treat it like an actual
-object, but both would lead to undefined behaviour of the transaction.
+object, but both would lead to undefined behavior of the transaction.
 
@@
 Degree 0 isolation to be the isolation level that doesn’t prohibit any abnormal
-behaviour. Furthermore, we define short duration locks as locks that are held
+behavior. Furthermore, we define short duration locks as locks that are held
 while a transaction reads or writes an item, in contrast to long-duration locks which are
 held until a transaction commits or aborts.
-These definition can then be used to clasify different database system implementations in
+These definition can then be used to classify different database system implementations in
 regards to their serializability.
@@
 \subsection{Repeatable read}
-For Repeatable read, the phenomenon P2, fuzzy read, has to be prohibited.
+For repeatable read, the phenomenon P2, fuzzy read, has to be prohibited.
 This requires long duration write locks as well as long duration read locks on
-items and short ones for Predicates.\\
+items and short ones for predicates.\\
 The difference to read committed is that read committed only guarantees that
@@
 on reads and writes. This isolation level is stronger then it's ANSI counter part due to the
-prohibition of prdicate updates and deletions as defined in P3, making it the only isolation level
-outside of the original ANSI ioslation level specifications.
+prohibition of predicate updates and deletions as defined in P3, making it the only isolation level
+outside of the original ANSI isolation level specifications.
 
@@
 \subsection{ P4 - lost update}
-This phenomenon describes the loss of an update by a transaction due to it imeadiatly being overwritten
+This phenomenon describes the loss of an update by a transaction due to it immediately being overwritten
 by a second transaction.
@@
     while the write lock only gets removed after the transaction commits or aborts.
-    As Cursor stability is stronger then read commited while being weaker then repeatable read it
-    is used instead of the weaker repeatable read by some database systems. This behaviour is allowed by the ANSI standard
+    As Cursor stability is stronger then read committed while being weaker then repeatable read it
+    is used instead of the weaker repeatable read by some database systems. This behavior is allowed by the ANSI standard
     as it defines a minimum set of guarantees.
 \end{example}
-\subsection{Stength}
-When looking at our isiolation levels defined in section 2, we can see that P4 is prohibited by repeatable read,
+\subsection{Strength}
+When looking at our isolation levels defined in section 2, we can see that P4 is prohibited by repeatable read,
 as long duration locks on read and write will also disallow the overwriting of an item by another transaction.
 It is thus weaker than repeatable read as prohibiting P4 doesn't prohibit P3.
-Looking at read commited one can reason that one transaction A can read X, then be interupted by transaction B which updates
+Looking at read committed one can reason that one transaction A can read X, then be interrupted by transaction B which updates
 item X and commits, and transaction A then continues to write to X.
-Thus makin Cursor stability stronger than the strngthened read commited.
+Thus making Cursor stability stronger than the strengthened read committed.
 \subsection{Multi cursor}
 While the Cursor stability based system described above could also be extended with the use of multiple cursors, thus
-parleying the effect of repeatable read isolation as long as each transaction doesn't acces more items
-then the system has cursors. But obiously this isn't a general, or even practical solution to the phenomenon P2 and
-although it might have real life usecases it won't be discussed further in this paper.
+parleying the effect of repeatable read isolation as long as each transaction doesn't access more items
+then the system has cursors. But obviously this isn't a general, or even practical solution to the phenomenon P2 and
+although it might have real life use cases it won't be discussed further in this paper.
 \section{Snapshot Isolation}
@@
 For the committing procedure, a second (commit) timestamp is assigned, said timestamp is larger than every
-other assigned time stamp in the system. If another transaction A already modified data of the commiting
+other assigned time stamp in the system. If another transaction A already modified data of the committing
 transaction B it's commit fails, leading to an abort.
 \subsection{P5B - Write skew}
-To take a closer look at the difference to repeatable read it is usefull to define a new phenomenon.
+To take a closer look at the difference to repeatable read it is useful to define a new phenomenon.
 
@@
 \begin{example}
-    (first commit wins): \\
-    Alice reads x and y to be 50€. Bob then reads the same data, while getting his own snapshot
-    to operate on (newer timestamp than Alice). Alice then sets y to be -40€ and commits,
-    resulting in a sum of 10€. Bob then sets y to be 30€ and tries to commit. As Alice already
-    modified x Bob’s transaction will abort.
+    (first commit wins):
+    Constrain: $ X+Y \geq  50$
+    Alice reads x to be 10€ and y to be 50€. Bob then reads the same data, while getting his own snapshot
+    to operate on (newer timestamp than Alice). Alice then sets x to be 5€,
+    resulting in a sum of 55€. Bob then sets y to be 40€. Alice commits and Bob then tries to commit.
 \end{example}
+This would be prohibited by repeatable read but is allowed in snapshot isolation.
+\subsection{Implications}
+When looking at the phenomena defined above one can see that prohibiting it with a locked based system
+would be the same as prohibiting P3 as a lost update can only be prohibited by placing long locks.
+But when using a multi version system instead proscribing P5B isn't equal to proscribing P3.
 
+The example in 5.1 is allowed by Snapshot isolation as it does not check for constrains and the transaction
+doesn't modify the same item (x and y respectively). Repeatable read on the other hand would disallow such
+behavior as it places long read and write locks on the items.
+P5B is thus useful to differentiate between Snapshot isolation and repeatable read.
 \subsection{Comparison with repeatable read}
-In comparison to snapshot isolation repeatable read prohibits P5B while snapshot isolation prohibits some cases of
-P3.
+As stated above repeatable read prohibits P5B while snapshot isolation prohibits some cases of
+P3 and both prohibit P0, P1, and P2. Although the argument above is made with a locking implementation
+it is general as the phenomena defined were strengthened to match the locking behavior.
 \subsection{Strength}
 Snapshot isolation is comparable in strength to repeatable read as both prohibit P0, P1, and P2.
-\subsection{Comparison to ANSI}
-When looking at the weak interpretation of the ANSI specification as in \cite{Adya_Liskov_O_Neil_2000}
-one can see that Snapshot isolation prohibits all three anomlies A1, A2, and A3 making it ANOMALY SERIALIZABLE.
-To show this we will now take a look at A3 as we have already shown above that Snapshot isolation prohibits P1 and P2.
+To further investigate their difference we have to take a look at the original ANSI specification of serializable,
+as it prohibits A3, which will be defined in the following section.
 \subsection{A3 - Anomaly Phantom}
-This is the strict interpretation of the Phatom anomaly. The difference to P3 is the position of the
-transactions comit is not indiferent.
+This is the strict interpretation of the Phantom anomaly. The difference to P3 is the position of the
+transactions commit is not indifferent for A3.
 \begin{example}
@@
 \end{example}
-Thus A3 only opperates on commited data while P3 accounts for transactions to still be in prosses.
+Thus A3 only operates on committed data while P3 accounts for transactions to still be in process,
+making A3 a weaker phenomenon than P3.
+\subsection{Comparison to ANSI}
+When looking at the weak interpretation of the ANSI specification as in \cite{Adya_Liskov_O_Neil_2000}
+one can see that Snapshot isolation prohibits all three anomalies A1, A2, and A3 making it ANOMALY SERIALIZABLE.
+To show this we will now take a look at A3 as we have already shown above that Snapshot isolation prohibits P1 and P2.
+As the definition of A3 states the importance of commit sequence it is easy to see that Snapshot isolation with a first commit wins
+strategy precludes A3, as a already altered item will lead to an abort.
+When looking at a first write wins the same is true as still only one modified value will
+be written to the database and the other one will be discarded by an abort. It thus only changes
+which transaction wins in comparison to first commit wins.
+\subsection{Comparison to Serializable}
+In comparison to the serializable isolation level defined in 3.4 we can say that Snapshot isolation is
+weaker as serializable precludes P5B as well as P3, which as shown in example 5.1 above doesn't hold true
+for Snapshot isolation.
 \section{Conclusion}
@@
 
-In conclusion one can see that the ANSI specification is inherintly flawed whwn talking
+In conclusion one can see that the ANSI specification is inherently flawed when talking
 about multi-version histories and non locking database systems, as it failed to consider
-these systems when defining the phenomena.
+these systems when defining the phenomena. As this will inevitably lead to different interpretations
+and implementations it might be advisable to update the ANSI standard
+according to the definitions given in \cite{Adya_Liskov_O_Neil_2000} and \cite{Berenson_Bernstein_Gray_Melton_O_Neil_O_Neil_1995}
+to ensure the correct behavior of non locking implementations of ANSI SQL.
 \bibliographystyle{plainnat}

### 2020-12-15 __latexindent_temp.tex
fixed minor errors

--- 2020-12-13 paper.tex
@@
 \author{Jeroen-Niclas Trzaska}
-\title{DB pro sem - isolation levels}
+\title{DB proseminar - Isolation Levels}
 \acmDOI{}
@@
 \begin{abstract}
-
     This paper is going to look at the strengthened ANSI definition propositions by \cite{Adya_Liskov_O_Neil_2000} as well as the corresponding isolation levels,
@@
 behavior of different database implementation approaches when presented with multi-item dependencies.
-This is important as the ANSI standard was created when locking implementations where the norm and thus multi version systems weren't
+This is important as the ANSI standard was created when locking implementations where the norm and thus multi version systems were not
 taken into account creating it leading to a behavior not in line with ANSI.
-These issues include but aren't limited to dependency constrains between two items.
+These issues include but are not limited to dependency constrains between two items.
 
-Let's say that Alice has two accounts at a bank. She has 20€ on each of them now she withdraws 30€ from account A and another 20€ form account B.
-The looking at each of the transactions we can see that the constrain $A+B \geq 0$ is met.
+For example assume that Alice has two accounts at a bank. She has 20€ on each of them. Now she withdraws 30€ from account A and another 20€ form account B.
+Looking at each of the transactions we can see that the constraint $A+B \geq 0$ is met.
 But when executed simultaneously the resulting sum can still be negative. To do so
@@
 \section{Phenomena Definitions}
-We're now going to introduce the strengthened ANSI phenomena definitions as described in \cite{Adya_Liskov_O_Neil_2000}  which
+We are now going to introduce the strengthened ANSI phenomena definitions as described in \cite{Adya_Liskov_O_Neil_2000}  which
 are based on the broad interpretation of the ANSI standard. These definitions are inherently stronger then their ANSI counter part,
-as they rule out other Phenomena not defined in the following paragraph. This is due to the fact that, as mentioned above, the ANSI standard
+as they rule out other Phenomena not defined in the following paragraph. This is due to the fact that the ANSI standard
 was created with locking implementations in mind. Leading to definitions that would allow phenomena to appear in the database history when using non lock based systems
-while lock based systems using the same interpretation wouldn't.
+while lock based systems using the same interpretation would not.
 
@@
 
-\includegraphics[width=8cm]{P0}
+\begin{figure}[h]
+
+\includegraphics[width=8.5cm]{P0}
+
+\end{figure}
 
@@
     To do so he reads the value of X and then adds 20€ and writes it back to the account X.
-    Alice than redacts her transaction.
+    Alice than performs a rollback on her transaction.
     Thus the amount of money in account X is unclear.
@@
 \subsection{P1 - Dirty read}
-For the second phenomenon we're going to look at reading inconsistent data.
+For the second phenomenon we are going to look at reading inconsistent data.
 This can be a problem when retrieving information while another transaction
@@
 
-\includegraphics[width=8cm]{P1}
+\begin{figure}[h]
+
+\includegraphics[width=8.5cm]{P1}
+
+\end{figure}
+
 \begin{example}
@@
 \subsection{P2 - Fuzzy read}
-The next phenomenon is similar to P1 as it to has to do with read inconsistencies.
+The next phenomenon is similar to P1 as it has to do with read inconsistencies, too.
 Though this time performing the two reads of the reading transaction without interruption would lead
-to the correct data, so the read is non repeatable. This is an issue as the transaction doesn't know that
-it was interrupted and assumes that the data in the database hasn't changed. Thus it operates under a false pretense
+to the correct data, so the read is non repeatable. This is an issue as the transaction does not know that
+it was interrupted and assumes that the data in the database has not changed. Thus it operates under a false pretense
 and might corrupt the database when writing data back to it.
 
-\includegraphics[width=8cm]{P2}
+\begin{figure}[h]
+
+    \includegraphics[width=8.5cm]{P2}
+
+\end{figure}
+
 \begin{example}
-    Bob starts to read the sum by reading x and gets interrupted by Alice who then transfers 40€ from x to y after this is completed Bob reads the value of y.
+    Bob starts to read the sum by reading x and gets interrupted by Alice who then transfers 40€ from x to y. 
+    After this is completed Bob reads the value of y.
     Thus the sum is 100€ for Alice and 140€ for Bob.
@@
 
-\includegraphics[width=8cm]{P3}
+\begin{figure}[h]
+
+    \includegraphics[width=8.5cm]{P3}
+
+\end{figure}
+
 \begin{example}
@@
 Not prohibiting this phenomenon can cause inconsistent predicate reads and in the worst case phantom entries in the
-database as the reading transaction could, after reading a phantom object, either assume that it doesn't exist or treat it like an actual
-object, but both would lead to undefined behavior of the transaction.
+database as the reading transaction could, after reading a phantom object, either assume that it does not exist or treat it like an actual
+object, both situations would lead to undefined behavior of the transaction.
 
@@
 We define the isolation levels by the phenomena prohibited by it and we define
-Degree 0 isolation to be the isolation level that doesn’t prohibit any abnormal
+Degree 0 isolation to be the isolation level that does not prohibit any abnormal
 behavior. Furthermore, we define short duration locks as locks that are held
@@
 We define read uncommitted to be the isolation level that prohibits the phenomenon
-P0, dirty write, this can be accomplished by placing long-duration write locks when modifying a data item.
+P0, dirty write. This can be accomplished by placing \emph{long-duration write locks} when modifying a data item.
 This only allows one transaction at a time to write to a given value and requires all others to wait for a commit or abort.
@@
 Read committed is defined as the isolation level where P1, dirty read, is prohibited.
-To accomplish this the use of short duration read locks, as well as long write locks, is
+To accomplish this the use of \emph{short duration read locks}, as well as long write locks, is
 necessary. Thus disallowing a write on data that was just read by another transaction as
@@
 For repeatable read, the phenomenon P2, fuzzy read, has to be prohibited.
-This requires long duration write locks as well as long duration read locks on
+This requires \emph{long duration write locks} as well as \emph{long duration read locks} on
 items and short ones for predicates.\\
@@
 \subsection{Serializable}
-For serializability P3 needs to be prohibited, which can be done by placing long locks
-on reads and writes. This isolation level is stronger then it's ANSI counter part due to the
+For serializability P3 needs to be prohibited, which can be done by placing \emph{long locks
+on reads and writes}. This isolation level is stronger then it's ANSI counter part due to the
 prohibition of predicate updates and deletions as defined in P3, making it the only isolation level
@@
 by a second transaction.
-\includegraphics[width=8cm]{P4}
+
+\begin{figure}[h]
+
+    \includegraphics[width=8.5cm]{P4}
+
+\end{figure}
+
 \begin{example}
     Alice reads x to be 100€. Bob then reads the same, adds 20€ and writes 120€ to x and commits.
-    Alice then subtracts 40€ from x and writes 60€ to x and commits. Resulting in the loss of Bob's update.
+    Alice then subtracts 40€ from x and writes 60€ to x and commits, resulting in the loss of Bob's update.
+\end{example}
 \subsection{Implications}
@@
     while the write lock only gets removed after the transaction commits or aborts.
-    As Cursor stability is stronger then read committed while being weaker then repeatable read it
+    As Cursor stability is stronger then read committed while being weaker then repeatable read, it
     is used instead of the weaker repeatable read by some database systems. This behavior is allowed by the ANSI standard
     as it defines a minimum set of guarantees.
-\end{example}
 \subsection{Strength}
@@
 as long duration locks on read and write will also disallow the overwriting of an item by another transaction.
-It is thus weaker than repeatable read as prohibiting P4 doesn't prohibit P3.
-Looking at read committed one can reason that one transaction A can read X, then be interrupted by transaction B which updates
+It is thus weaker than repeatable read as prohibiting P4 does not prohibit P3.
+Looking at read committed, one can reason that one transaction A can read X, then be interrupted by transaction B which updates
 item X and commits, and transaction A then continues to write to X.
@@
 While the Cursor stability based system described above could also be extended with the use of multiple cursors, thus
-parleying the effect of repeatable read isolation as long as each transaction doesn't access more items
-then the system has cursors. But obviously this isn't a general, or even practical solution to the phenomenon P2 and
-although it might have real life use cases it won't be discussed further in this paper.
+parleying the effect of repeatable read isolation as long as each transaction does not access more items
+then the system has cursors. But obviously this is not a general, or even practical solution to the phenomenon P2 and
+although it might have real life use cases it wo not be discussed further in this paper.
 \section{Snapshot Isolation}
-All of the above-mentioned systems suggest lock-based implementations, in contrast we will now look at
+All of the above-mentioned systems suggest lock-based implementations. In contrast we will also look at
 snapshot isolation which uses multi-versioning.
@@
 
-\includegraphics[width=8cm]{P5}
+\begin{figure}[h]
+
+    \includegraphics[width=8.5cm]{P5}
+
+    \end{figure}
 \begin{example}
     (first commit wins):
-    Constrain: $ X+Y \geq  50$
+    Constraint: $ X+Y \geq  50$
     Alice reads x to be 10€ and y to be 50€. Bob then reads the same data, while getting his own snapshot
@@
 would be the same as prohibiting P3 as a lost update can only be prohibited by placing long locks.
-But when using a multi version system instead proscribing P5B isn't equal to proscribing P3.
+But when using a multi version system instead proscribing P5B is not equal to proscribing P3.
 
-The example in 5.1 is allowed by Snapshot isolation as it does not check for constrains and the transaction
-doesn't modify the same item (x and y respectively). Repeatable read on the other hand would disallow such
+The example in subsection 5.1 is allowed by Snapshot isolation as it does not check for constrains and the transaction
+does not modify the same item (x and y respectively). Repeatable read on the other hand would disallow such
 behavior as it places long read and write locks on the items.
@@
 \subsection{A3 - Anomaly Phantom}
-This is the strict interpretation of the Phantom anomaly. The difference to P3 is the position of the
+This is the strict interpretation of the Phantom anomaly. The difference to P3 is that the position of the
 transactions commit is not indifferent for A3.
@@
 \end{example}
-Thus A3 only operates on committed data while P3 accounts for transactions to still be in process,
+Thus A3 only operates on committed data while P3 accounts for transactions to still be in progress,
 making A3 a weaker phenomenon than P3.
@@
 \subsection{Comparison to Serializable}
-In comparison to the serializable isolation level defined in 3.4 we can say that Snapshot isolation is
-weaker as serializable precludes P5B as well as P3, which as shown in example 5.1 above doesn't hold true
+In comparison to the serializable isolation level defined in section 3.4 we can say that Snapshot isolation is
+weaker as serializable precludes P5B as well as P3, which as shown in example 5.1 above does not hold true
 for Snapshot isolation.

### 2020-12-20 paper.tex
rc_1 update of paper

--- 2020-12-13 paper.tex
@@
-\documentclass[sigconf, review=true]{acmart}
+\documentclass[sigconf, review=false]{acmart}
 \usepackage[english]{babel}
@@
 \author{Jeroen-Niclas Trzaska}
-\title{DB pro sem - isolation levels}
+\title{DB proseminar - Isolation Levels}
 \acmDOI{}
 \acmISBN{}
+\acmConference{Proseminar}{WS2020/2021}{}
+\settopmatter{printacmref=false}
+
 \affiliation{%
@@
    \country{Germany}}
-\email{jeroen@trzaska.xyz}
+\email{jeroen-niclas.trzaska@mailbox.tu-dresden.de}
 
@@
 \begin{abstract}
-
-    This paper is going to look at the strengthend ANSI definition poropositions by \cite{Adya_Liskov_O_Neil_2000} as well as the coresponding isolation levels,
-    comparing them in terms of strength. We will also take a quick look at non lock based implementations
-    and the problems the ANSI standard has in regards to such implementations as shown in \cite{Berenson_Bernstein_Gray_Melton_O_Neil_O_Neil_1995}.
+    This paper is going to look at the strengthened ANSI SQL phenomena definition propositions
+    by \cite{Adya_Liskov_O_Neil_2000} as well as the corresponding isolation levels,
+    comparing them in terms of strength. It will also take a quick look at non lock based implementations
+    and the problems the ANSI SQL standard has in regards to such implementations as shown in \cite{Berenson_Bernstein_Gray_Melton_O_Neil_O_Neil_1995}.
 \end{abstract}
@@
 \section{Motivation}
-The motivation behind these analyses of the ANSI standard is to better define and examine the
-behaviour of different database implementation aproaches when presented with multi-item dependencies.
-This is important as the ANSI standard was created when locking implementations where the norm and thus multi version systems weren't
-taken into account creating it leading to a behaviour not in line with ANSI.
-These issues include but aren't limited to dependency constrains between two items.
-
-Let's say that Alice has two acoounts at a bank. She has 20€ on each of them now she whithdraws 30€ from acount A and another 20€ form account B.
-The looking at each of the transactions we can see that the constrain $A+B \geq 0$ is met.
-But when executed simultaniously the resulting sum can still be negative. To do so
-she asked Bob to help her.
+The motivation behind these analyses of the ANSI SQL standard is to better define and examine the
+behavior of different database implementation approaches when presented with multi-item dependencies.
+This is important as the ANSI standard was created when locking implementations where the norm and thus multi version systems were not
+taken into consideration when creating it, leading to unexpected behavior not in line with ANSI SQL.
+These issues include but are not limited to dependency constrains between two items in a database. \\
+For example assume that Alice has two accounts at a bank, the accounts can be overdrawn as long as the sum is greater or equal to 0.
+She has 20€ on each of the accounts. Now she withdraws 30€ from account A and another 20€ form account B.
+Looking at each of the transactions out of context, we can see that the constraint $A+B \geq 0$ is met.
+But when executed simultaneously the resulting sum of $A$ and $B$ can still be negative. To achieve this outcome
+Alice asks Bob to help her.
 \begin{example}
-    Alice reads 20€ to be in both accounts and Bob does the same. Alice then withdraws 30€ from account A and writes -10€ into it and commits, the constrain is still met.
-    Bob then withdraws 20€ from account and, as his transaction read both acounts to have 20€, he writes 0€  to B. This now violates the constrain $A+B \geq 0$.
+    Alice reads 20€ to be in both accounts and Bob does the same. Alice then
+    withdraws 30€ from account A and writes -10€ into it and commits, the constrain,
+    $A+B \geq 0$, is still met.
+    Bob then withdraws 20€ from account and, as his transaction had read both accounts to have
+    20€, he writes 0€  to B. Both transactions commit and the committed data in the database now violates the constrain $A+B \geq 0$.
 \end{example}
-\section{Phemomena Definitions}
-We're now going to introduce the strengthened ANSI phenomena definitions as described in \cite{Adya_Liskov_O_Neil_2000}  which
-are based on the broad interpretation of the ANSI standard. These definitions are inherinty stronger then their ANSI counter part,
-as they rule out other Phenomena not defined in the following paragraph. This is due to the fact that, as mentioned above, the ANSI standard
-was created with locking implementations in mind. Leading to definitions that would allow phenomena to apear in the database history when using non lock based systems
-while lock based systems using the same interpretation wouldn't.
+The following figure visualizes the relation between the phenomena and isolation levels that will be defined
+at a later point in this paper. Starting with the weakest level of isolation (Degree 0) each arrow denotes the phenomenon to be prohibited
+by the next isolation level.
+\begin{figure}[H]
+    \includegraphics[width=8.5cm, height=9.5cm]{iso_lvl_dia.png}
+    \caption{isolation level diagram}
+    \Description{isolation level diagram}
+\end{figure}
+\section{Phenomena Definitions}
+We are now going to introduce the strengthened ANSI phenomena definitions as described in \cite{Adya_Liskov_O_Neil_2000} which
+are based on the broad interpretation of the ANSI SQL standard. These definitions are inherently stronger then their ANSI counter part,
+as they rule out more phenomena as defined in the following paragraphs. This is due to the fact that the ANSI standard
+was created with locking implementations in mind. Leading to definitions that would allow phenomena to appear in the database's history when using non lock based systems
+while lock based systems using the same interpretation would not allow them.
 
 \subsection{P0 - Dirty write}
-An example for the above mentioned problem with ANSI strength is that while a locking implementation of read uncommitted
-prevents the following phenomenon, a non locking system would not do so.
-So lets start by defining the first phenomena, which describes two consecutive writes by two differnt
-transactions leading to an unclear value for X when the first transaction aborts.
-
-\includegraphics[width=8cm]{P0}
+An example for the above mentioned problem with the strength of ANSI SQL is that while a locking implementation of \emph{read uncommitted}
+prevents the following phenomenon, a non locking system would not do so (see s
+section 2.2 of \cite{Adya_Liskov_O_Neil_2000}).
+We start by defining the first phenomena, which describes two consecutive writes by two different
+transactions leading to an undefined value for x in case the first transaction aborts.
 
 \begin{example}
-    Alice deposits 10€ into a bank account X.
+    Alice deposits 10€ into a bank account x.
     Bob than deposits a further 20€ into the same account.
-    To do so he reads the value of X and then adds 20€ and writes it back to the account X.
-    Allice than redacts her transaction.
-    Thus the amount of monney in account X is unclear.
+    To do so he reads the value of x and then adds 20€ and writes it back to the account x.
+    Alice than performs a rollback on her transaction.
+    Thus the amount of money in account x is unclear.
 \end{example}
 
+\begin{figure}[h]
+    \includegraphics[width=8.5cm]{P0}
+    \caption{History for P0}
+    \Description{P0}
+\end{figure}
+
 \subsection{P1 - Dirty read}
-For the second phenomenon we're going to look at reading incosistent data.
+For the second phenomenon we are going to look at reading inconsistent data from the database.
 This can be a problem when retrieving information while another transaction
-that writes data is being executed. An issue that can arive due to this phenomena is
-that the given transaction reading the inconsistent data might continue to work with this
+that writes data is being executed. An issue that can arise when this phenomenon is not prohibited,
+is that the given transaction reading the inconsistent data will continue to work with this
 incorrect data and writing wrong data back to the database, corrupting it in the process.
 
-\includegraphics[width=8cm]{P1}
+\begin{figure}[h]
+    \includegraphics[width=8.5cm]{P1}
+    \caption{History for P1}
+    \Description{P1}
+\end{figure}
+
 \begin{example}
     Alice transfers 40€ form account x to account y.
-    To do so she reads the value of x subtracts the 40€ writing 10€ to value x.
-    Afterwards, Bob reads the value of x (10€) and y(15€) resulting in a sum of 25€
-    Alice then reads the value of y (15€) and then adds the 40€ to it resulting in a 55€ sum for y
+    To do so she reads the value of x subtracts the 40€ writing 10€ to value of x.
+    Afterwards, Bob reads the value of x (10€) and y (15€) resulting in a sum of 25€
+    Alice then reads the value of y (15€) and then adds the 40€ to it resulting in a sum of 55€ for y
     and a total of 65€.
@@
 \subsection{P2 - Fuzzy read}
-The next phenomenon is similar to P1 as it to has to do with read inconsitencies.
-Though this time performing the two reads of the reading transaction without interuption would lead
-to the correct data, so the read is non repeatable. This is an issue as the transaction doesn't know that
-it was interupted and assumes that the data in the database hasn't changed. Thus it opperates under a false pretense
+The next phenomenon is similar to \textbf{P1} as it has to do with read inconsistencies, too.
+Though this time performing the two reads of the reading transaction without interruption would lead
+to the correct data, so we call this read non repeatable. This is an issue as the transaction does not know that
+it got interrupted and assumes that the data in the database has not changed. Thus it operates under a false assumption
 and might corrupt the database when writing data back to it.
 
-\includegraphics[width=8cm]{P2}
+\begin{figure}[h]
+    \includegraphics[width=8.5cm]{P2}
+    \caption{History for P2}
+    \Description{P2}
+\end{figure}
+
 \begin{example}
-    Bob starts to read the sum by reading x and gets interrupted by Alice who then transfers 40€ from x to y after this is completed Bob reads the value of y.
-    Thus the sum is 100€ for Alice and 140€ for Bob.
-    In contrast to P1 a second read of x by Bob would have returned the correct sum.
+    Bob starts to read the sum by reading x and gets interrupted by Alice who transfers 40€ from account x to y.
+    After this is completed Bob reads the value of y.
+    Thus the sum is 100€ for Alice while it is 140€ for Bob.
+    In contrast to \textbf{P1} a second read of x by Bob would have returned the correct sum of 100€.
 \end{example}
 \subsection{P3 - Phantom}
-Once again when looking at the ANSI specification one can see that it only disallowes inserts on the serializable
-isolation level while the following definition of P3 prohibits deletions, inserts, and updates, making it stronger than A3.
-This phonomenon is only applicable to transactions using predicates and aplies to deletions, insets, and upadates.
+Once again when looking at the ANSI specification one can see that it only disallows inserts on the \emph{serializable}
+isolation level while the following definition of \textbf{P3} prohibits deletions, inserts, and updates, making it stronger than \textbf{A3}.
+This phenomenon is only applicable to transactions using predicates and applies to deletions, insets, and updates.
 
-\includegraphics[width=8cm]{P3}
+\begin{figure}[h]
+    \includegraphics[width=8.5cm]{P3}
+    \caption{History for P3}
+    \Description{P3}
+\end{figure}
+
 \begin{example}
@@
 Not prohibiting this phenomenon can cause inconsistent predicate reads and in the worst case phantom entries in the
-database as the reading transaction could, after reading a phantom object, either assume that it doesn't exist or treat it like an actual
-object, but both would lead to undefined behaviour of the transaction.
+database as the reading transaction could, after reading a phantom object, either assume that it does not exist or treat it like an actual
+object,. Both situations would lead to undefined behavior of the transaction.
 
@@
 We define the isolation levels by the phenomena prohibited by it and we define
-Degree 0 isolation to be the isolation level that doesn’t prohibit any abnormal
-behaviour. Furthermore, we define short duration locks as locks that are held
+Degree 0 isolation to be the isolation level that does not prohibit any abnormal
+behavior. Furthermore, we define short duration locks as locks that are held
 while a transaction reads or writes an item, in contrast to long-duration locks which are
 held until a transaction commits or aborts.
-These definition can then be used to clasify different database system implementations in
+These definition can then be used to classify different database system implementations in
 regards to their serializability.
@@
 \subsection{Read uncommitted}
-We define read uncommitted to be the isolation level that prohibits the phenomenon
-P0, dirty write, this can be accomplished by placing long-duration write locks when modifying a data item.
-Thus only allowing one transaction at a time to modify a given value.
+We define \emph{read uncommitted} to be the isolation level that prohibits the phenomenon
+P0, \emph{dirty write}. This can be accomplished by placing \emph{long-duration write locks} when modifying a data item.
+This only allows one transaction at a time to write to a given value and requires all others to wait for a commit or abort.
 
 \subsection{Read committed}
-Read committed is defined as the isolation level where P1, dirty read, is prohibited.
-To accomplish this the use of short duration read locks, as well as long write locks, is
-necessary. Thus disallowing a write on data that was just read by another transaction as
-well as disallowing overwriting of an item already written by another, not finished, transaction.
+Read committed is defined as the isolation level where \textbf{P1}, \emph{dirty read} is prohibited.
+To accomplish this the use of \emph{short duration read locks}, as well as \emph{long write locks}, is
+necessary. Thus disallowing a write on an data item that was just read by another transaction, as
+well as disallowing overwriting of an item already written by another, unfinished, transaction.
 
 \subsection{Repeatable read}
-For Repeatable read, the phenomenon P2, fuzzy read, has to be prohibited.
-This requires long duration write locks as well as long duration read locks on
-items and short ones for Predicates.\\
-The difference to read committed is that read committed only guarantees that
-the data read was committed at the time of reading (no dirty reads), whereas
-repeatable read also guarantees that the data will not change before the
+For \emph{repeatable read}, the phenomenon \textbf{P2}, \emph{fuzzy read}, has to be prohibited.
+This requires \emph{long duration write locks} as well as \emph{long duration read locks} on
+items and \emph{short duration locks} on predicates.\\
+The difference to \emph{read committed} is that in the \emph{read committed} isolation level it is only guaranteed that
+the data read was committed at the time of reading (no dirty reads). Whereas
+\emph{repeatable read} also guarantees that the data read will not change before the
 transaction finishes either with a commit or abort.
@@
 \subsection{Serializable}
-For serializability P3 needs to be prohibited, which can be done by placing long locks
-on reads and writes. This isolation level is stronger then it's ANSI counter part due to the
-prohibition of prdicate updates and deletions as defined in P3, making it the only isolation level
-outside of the original ANSI ioslation level specifications.
+For serializability the phenomenon \textbf{P3} needs to be prohibited, which can be done by placing \emph{long locks
+    on reads and writes}. This isolation level is stronger then it's ANSI counter part due to the
+prohibition of predicate updates and deletions as defined in \textbf{P3}, making it the only isolation level
+stronger than the original ANSI serializable isolation level specification.
 
 \section{Cursor stability}
-For the cursor stability, it is useful to define a fourth phenomenon P4 (lost update), as this isolation level
+For the isolation level \emph{cursor stability}, it is useful to define a fourth phenomenon \textbf{P4} (\emph{lost update}), as this isolation level
 was designed to solve the problem of lost updates.
 \subsection{ P4 - lost update}
-This phenomenon describes the loss of an update by a transaction due to it imeadiatly being overwritten
+This phenomenon describes the loss of an update to an data item by a transaction, due to it immediately being overwritten
 by a second transaction.
-\includegraphics[width=8cm]{P4}
+
+\begin{figure}[h]
+    \includegraphics[width=8.5cm]{P4}
+    \caption{History for P4}
+    \Description{P4}
+\end{figure}
+
 \begin{example}
     Alice reads x to be 100€. Bob then reads the same, adds 20€ and writes 120€ to x and commits.
-    Alice then subtracts 40€ from x and writes 60€ to x and commits. Resulting in the loss of Bob's update.
+    Alice then subtracts 40€ from x and writes 60€ to x and commits, resulting in the loss of Bob's update.
+\end{example}
+
 \subsection{Implications}
-    The phenomena define above can be prohibited when a read lock is placed on x (position of the cursor) when
+The phenomena define above can be prohibited by placing a read lock on x (position of the cursor) when
 read and placing a long write lock on the items row when modified.
-    The read lock is removed when the cursor moves (e.g. reading a new cell),
-    while the write lock only gets removed after the transaction commits or aborts.
-    As Cursor stability is stronger then read commited while being weaker then repeatable read it
-    is used instead of the weaker repeatable read by some database systems. This behaviour is allowed by the ANSI standard
+The read lock is removed when the cursor moves (e.g. reading a new tuple),
+while the write lock only gets removed after the transaction commits or aborts.\\
+As \emph{Cursor stability} is stronger then \emph{read committed} while being weaker then \emph{repeatable read}, it
+is used instead of the weaker \emph{repeatable read} by some database systems. This behavior is allowed by the ANSI standard
 as it defines a minimum set of guarantees.
-\end{example}
-\subsection{Stength}
-When looking at our isiolation levels defined in section 2, we can see that P4 is prohibited by repeatable read,
-as long duration locks on read and write will also disallow the overwriting of an item by another transaction.
-It is thus weaker than repeatable read as prohibiting P4 doesn't prohibit P3.
-Looking at read commited one can reason that one transaction A can read X, then be interupted by transaction B which updates
-item X and commits, and transaction A then continues to write to X.
-Thus makin Cursor stability stronger than the strngthened read commited.
+\subsection{Strength}
+When looking at the isolation levels defined in section 2, one can see that \textbf{P4} is prohibited by \emph{repeatable read},
+as \emph{long duration locks} on reads and writes will also disallow the overwriting of an item by another transaction.
+\emph{cursor Stability} is thus weaker than \emph{repeatable read} as prohibiting \textbf{P4} does not prohibit \textbf{P3}.
+Looking at \emph{read committed}, one can reason that a transaction A can read x, then be interrupted by transaction B which updates
+item x and commits, and transaction A then continues to write to x.
+Thus making \emph{Cursor stability} stronger than the strengthened \emph{read committed}.
 \subsection{Multi cursor}
-While the Cursor stability based system described above could also be extended with the use of multiple cursors, thus
-parleying the effect of repeatable read isolation as long as each transaction doesn't acces more items
-then the system has cursors. But obiously this isn't a general, or even practical solution to the phenomenon P2 and
-although it might have real life usecases it won't be discussed further in this paper.
+While the \emph{Cursor stability} based system described above could also be extended with the use of multiple cursors, thus
+parleying the effect of \emph{repeatable read} isolation, as long as each transaction does not access more items
+then the system has cursors. But obviously this is neither a general, nor even a practical solution to the phenomenon \textbf{P2} and
+although it might have real life use cases it will not be discussed further in this paper.
 \section{Snapshot Isolation}
-All of the above-mentioned systems suggest lock-based implementations, in contrast we will now look at
-snapshot isolation which uses multi-versioning.
-In snapshot isolation, each transaction gets its own snapshot of committed data at the start.
-Thus reads are never blocked, while writes are performed on the given snapshot as to not disturb
-other transactions and allow reads of the updated data only by the same transaction (unless committed).
+While all of the above-mentioned systems suggest lock-based implementations, this is not the only way to prohibit the defined phenomena.
+In contrast we will look at
+\emph{snapshot isolation} which uses multi-versioning to prohibit these phenomena.\\
+In \emph{snapshot isolation}, each transaction gets it's own snapshot of all committed data at it's start,
+thus reads are never blocked. All writes are performed on the given snapshot as to not disturb
+other transactions. Reads of the updated data is only allowed for the same transaction that wrote them (unless the transaction has committed).
 For the committing procedure, a second (commit) timestamp is assigned, said timestamp is larger than every
-other assigned time stamp in the system. If another transaction A already modified data of the commiting
+other assigned time stamp in the system. If another transaction A already modified data of the committing
 transaction B it's commit fails, leading to an abort.
 \subsection{P5B - Write skew}
-To take a closer look at the difference to repeatable read it is usefull to define a new phenomenon.
+To take a closer look at the difference to \emph{repeatable read} it is useful to define a new phenomenon.
 
-\includegraphics[width=8cm]{P5}
+\begin{figure}[h]
+    \includegraphics[width=8.5cm]{P5}
+    \caption{History for P5}
+    \Description{P5}
+\end{figure}
+
 \begin{example}
-    (first commit wins): \\
-    Alice reads x and y to be 50€. Bob then reads the same data, while getting his own snapshot
-    to operate on (newer timestamp than Alice). Alice then sets y to be -40€ and commits,
-    resulting in a sum of 10€. Bob then sets y to be 30€ and tries to commit. As Alice already
-    modified x Bob’s transaction will abort.
+    (assuming first commit wins):
+    Constraint: $ x+y \geq  50$
+    Alice reads x to be 10€ and y to be 50€. Bob then reads the same data, while getting his own snapshot
+    to operate on (newer timestamp than Alice). Alice then sets x to be 5€,
+    resulting in a sum of 55€. Bob then sets y to be 40€. Alice commits and Bob then commits resulting in a sum of 45€.
 \end{example}
 
+\subsection{Implications}
+When looking at the phenomena defined above one can see that prohibiting it with a locked based system
+would be equal to prohibiting \textbf{P3} as a lost update can only be prohibited by placing long locks on all reads and writes.
+But when using a multi version system instead, proscribing \textbf{ \textbf{P5B}} is not equal to proscribing \textbf{P3}.\\
+The example in subsection 5.1 is allowed by \emph{Snapshot isolation} as it does not check for constrains and the transactions
+does not modify the same item (x and y respectively). \emph{Repeatable read} on the other hand would disallow such
+behavior as it places \emph{long read} and \emph{write locks} on the items.\\
+\textbf{P5B} is thus useful to differentiate between \emph{Snapshot isolation} and \emph{repeatable read}.
 \subsection{Comparison with repeatable read}
-In comparison to snapshot isolation repeatable read prohibits P5B while snapshot isolation prohibits some cases of
-P3.
-\subsection{Strength}
-Snapshot isolation is comparable in strength to repeatable read as both prohibit P0, P1, and P2.
-\subsection{Comparison to ANSI}
-When looking at the weak interpretation of the ANSI specification as in \cite{Adya_Liskov_O_Neil_2000}
-one can see that Snapshot isolation prohibits all three anomlies A1, A2, and A3 making it ANOMALY SERIALIZABLE.
-To show this we will now take a look at A3 as we have already shown above that Snapshot isolation prohibits P1 and P2.
+As stated above \emph{repeatable read} prohibits \textbf{P5B} while \emph{snapshot isolation} prohibits some cases of
+\textbf{P3} and both prohibit \textbf{P0}, \textbf{P1}, and \textbf{P2}. Although the argument above is made with a locking implementation
+it is general as the phenomena defined were strengthened to match the behavior of locking implementations. \\
+\emph{Snapshot isolation} is comparable in strength to \emph{repeatable read} as both prohibit \textbf{P0}, \textbf{P1}, and \textbf{P2}.
+To further investigate their difference we have to take a look at the original ANSI specification of \emph{serializable},
+as it prohibits	\textbf{A3}, which will be defined in the following section.
 \subsection{A3 - Anomaly Phantom}
-This is the strict interpretation of the Phatom anomaly. The difference to P3 is the position of the
-transactions comit is not indiferent.
+This is the strict interpretation of the Phantom anomaly. The difference to \textbf{P3} is that the position of the
+transactions commit is not indifferent for \textbf{A3}.
 \begin{example}
@@
 \end{example}
-Thus A3 only opperates on commited data while P3 accounts for transactions to still be in prosses.
+Thus \textbf{A3} only operates on committed data while \textbf{P3} also accounts for transactions to still be in progress,
+making \textbf{A3} a weaker phenomenon than \textbf{P3}.
+\subsection{Comparison to ANSI}
+When looking at the weak interpretation of the ANSI specification as in \cite{Adya_Liskov_O_Neil_2000}
+one can see that Snapshot isolation prohibits all three anomalies \textbf{P1}, \textbf{P2}, and \textbf{A3} making it \emph{anomaly \emph{serializable}}.
+To show this we will now take a look at \textbf{A3} as we have already shown above that Snapshot isolation prohibits \textbf{P1} and \textbf{P2}.
+As the definition of \textbf{A3} states the importance of commit sequence it is easy to see that Snapshot isolation with a first commit wins
+strategy precludes \textbf{A3}, as a already altered item will lead to an abort of the transaction that tries to commit.
+When looking at a first write wins the same is true as still only one modified value will
+be written to the database and the other one will be discarded by an abort. It thus only changes
+which transaction wins in comparison to first commit wins.
+\subsection{Comparison to Serializable}
+In comparison to the \emph{serializable} isolation level defined in section 3.4 we can say that Snapshot isolation is
+weaker as \emph{serializable} precludes \textbf{P5B} as well as \textbf{P3}, which as shown in example 5.1 above does not hold true
+for Snapshot isolation.
 \section{Conclusion}
-\includegraphics[width=8cm, height=8cm]{iso_lvl_dia.png}
 
-In conclusion one can see that the ANSI specification is inherintly flawed whwn talking
-about multi-version histories and non locking database systems, as it failed to consider
-these systems when defining the phenomena.
+In conclusion one can see that the ANSI SQL specification is inherently flawed when talking
+about multi-version histories and other non locking database systems, as it failed to consider
+these systems when defining the phenomena. This will inevitably lead to different interpretations
+and implementations. To ensure the correct behavior of non locking implementations of ANSI SQL it might be advisable to update the ANSI standard
+according to the definitions discussed above and in \cite{Adya_Liskov_O_Neil_2000} and \cite{Berenson_Bernstein_Gray_Melton_O_Neil_O_Neil_1995}.
 \bibliographystyle{plainnat}
